---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# paindrawr

<!-- badges: start -->
<!-- badges: end -->

**paindrawr** helps you import, transform, quantify, and visualise *pain
drawings* — schematic illustrations on which patients mark where they feel
pain on an outline of the human body.

Pain drawings have traditionally been scored by dividing the body into a
fixed set of anatomical regions and recording, for each region, whether it
was marked or not. Modern digital tools (web and tablet apps) instead capture
each drawing as a series of pen-stroke *x/y coordinates*. This coordinate-based
data is far richer, but it is also harder to work with. paindrawr provides a
consistent workflow for turning that raw coordinate data into figures and
quantitative summaries.

## Installation

You can install the development version of paindrawr from
[GitHub](https://github.com/steenharsted/paindrawr) with:

```{r}
#| eval: false
pak::pak("steenharsted/paindrawr")
```


## Data structure

paindrawr works with a tibble in which each row is one respondent. The pain
drawings themselves live in a dedicated list-column, so that each
drawing — its metadata, strokes, and point coordinates — travels alongside any
covariates you have collected (for example sex, age, symptom duration, or pain
intensity). The list column name defaults to `pdr_data`, but can take any name. 
The package ships with `pdr_example_tibble`, a ready-to-use example
in exactly this format, which is used throughout the documentation.

A detailed description of the pain-drawing data structure will be provided in a
separate vignette.

## Example

paindrawr is built to work within the
[tidyverse](https://www.tidyverse.org/): drawings are stored in a list-column
of a tibble, and the `pdr_*()` functions are designed to be used with the
native pipe (`|>`) alongside familiar verbs like `mutate()`, `filter()`, and
`head()`. The examples below load the tidyverse for this reason.

The example dataset contains 50 respondents, each with a pain drawing in the
`pdr_data` column plus a few covariates:

```{r example}
library(paindrawr)
library(tidyverse)

pdr_example_tibble
```

Metadata stored inside each drawing can be pulled out with `pdr_get_info()`.
Because the drawings live in a list-column, you can add drawing-level metadata
as new columns with `mutate()` — for example the canvas width of each drawing:

```{r get-info}
pdr_example_tibble |>
  mutate(width = pdr_get_info(pdr_data, var = ".width"))
```

### Recreate a single drawing

`pdr_plot_drawing()` redraws an individual pain drawing from its coordinates.
Pipe in a single row of the tibble:

```{r plot-drawing}
pdr_example_tibble |>
  head(1) |>
  pdr_plot_drawing()
```

### Heatmaps across groups

`pdr_plot_heatmap()` aggregates many drawings into density heatmaps, optionally
stratified by one or two grouping variables. Here we stratify by symptom
duration:

```{r plot-heatmap, message = FALSE}
pdr_example_tibble |>
  pdr_plot_heatmap(variables = "duration")
```



## Typical workflow

paindrawr functions are organised around a common analysis pipeline:

- **Import** raw drawings — `pdr_import_json()`, `pdr_import_mird()`
- **Check** that the data is well-formed — `pdr_check_data()`
- **Modify** and clean drawings — `pdr_modify()`, `pdr_anatomy_merge()`
- **Quantify** the marked areas — `pdr_poly_areas()`, `pdr_get_alpha_area()`
- **Visualise** the results — `pdr_plot_drawing()`, `pdr_plot_heatmap()`

## Rendering this README

This file is generated from `README.Rmd`. After editing, re-render it with
`devtools::build_readme()` and commit the resulting `README.md` and any figure
files under `man/figures/` so they display on GitHub.
